### 2016-07-27 capability-machine.tex
Updated malloc specification

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments\ldotp\\
    &\quad \forall n, i, \size \in \nats\ldotp\forall b,e,a,m_b,m_e\in \Addrs\ldotp\forall p \in \Perms \ldotp\\
    &\qquad W_{\malloc} = [i \mapsto \iota_{\malloc}] \land \\
    &\qquad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \heapSat[\hsfoot]{n}{W_{\malloc}} \land \\
    &\qquad \memreg(r_1) = \size \land \\
    &\qquad \size \geq 0 \land \\
    &\qquad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad\Rightarrow \\
    &\qquad \quad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
    &\qquad \qquad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \\
    &\qquad \qquad \quad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \quad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
    &\qquad \qquad \quad W' = [i \mapsto \iota_{\malloc}, k \mapsto \iota^0_{b',e'}] \land \\
    &\qquad \qquad \quad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \quad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \quad \size = e'-b'
    &\qquad \qquad \quad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \quad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
  pc unspecified as it is not important for malloc users?}
\dominique{Resulting frame should be equal to original.}
\dominique{Allocated pointer should have the right size.}
\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}

@@

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2 \in \phi_1)
\]
Heap segment satisfaction:
\[
\heapSat[\hs]{n}{W} \text{ iff }
\begin{aligned}
  \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{aligned}
\]
Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.

A standard heap invariant that insures all values in the region are in the value  relation:
\[
  \iota_{\var{start},\var{end}} \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{start},\var{end}}(W) & \defeq (s,=,H_{\var{std}}) \\
  H_{\var{std}}\; s\; W & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(a) \in \stdvr(\roll \; W)\}
\end{align*}
Another standard heap invariant that ensures all the words in the region are 0.
\[
  \iota_{\var{start},\var{end}}^0 \; : \; \StorePred
\]
\begin{align*}
  \iota_{\var{start},\var{end}}^0(W) \defeq \{ \npair{\heapseg} \mid & \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
\end{align*}

@@

\begin{definition}[Heap satisfaction/erasure]
\begin{align*}
  \heapSat{n}{W} & & \text{iff} & &
                                      \arraycolsep=0pt
                                      \begin{array}{l}
                                      \exists r: \dom(W) \rightarrow \HeapSegments \ldotp \\
                                      \qquad\heap = \biguplus_{\iota\in \dom(W)} r(\iota)\\
                                      \quad\text{and}\\
                                      \qquad\forall \iota \in \dom(W) \ldotp \npair{r(\iota)} \in W(\iota)(W)
                                      \end{array}\\
\end{align*}

\end{definition}
\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  For any $n \in \nats$, $W \in \Worlds$, $p\in \Perms$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction. Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
\lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR. What should these assumptions be? So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition from this.}
\dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}

>>>>

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \var{FS} \ldotp \\
    &\qquad \quad W (i) = \iota_{\malloc}^{\var{FS}} \land \\
    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' \future W \land \\
    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\var{FS} \setminus [b',e']} \land \\
    &\qquad \qquad \qquad [b',e'] \subseteq \var{FS}  \land \\
    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \qquad \size = e'-b' \\
    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
  pc unspecified as it is not important for malloc users?}
\dominique{Resulting frame should be equal to original.}
\dominique{Allocated pointer should have the right size.}
\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}

@@

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2 \in \phi_1)
\]
\begin{definition}[Heap satisfaction/erasure]
\[
\heapSat[\hs]{n}{W} \text{ iff }
\begin{gathered}
  \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}\\
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{gathered}
\]
Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.
\end{definition}

A standard heap invariant that insures all values in the region are in the value  relation:
\[
  \iota_{\var{start},\var{end}} \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{base},\var{end}}(W) & \defeq (s,=,H_{\var{base},\var{end}}) \\
  H_{\var{base},\var{end}}\; s\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
                             & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
\end{align*}
Another standard heap invariant that ensures all the words in the region are 0.
\[
  \iota_{\var{start},\var{end}}^0 \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq (s,=,H_{\var{base},\var{end}}^0) \\
  H_{\var{base},\var{end}}^0\; s\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
\end{align*}

@@

\begin{lemma}[Fundamental theorem of logical relations (FTLR)] \forcenewline
  For any $n \in \nats$, $W \in \Worlds$, $p\in \Perms$, and addresses $\var{base}$, $\var{end}$, and $a$, if $\perm = \rwx$ and $\writeCond{n,W,\var{base},\var{end}}$ or $\perm = \exec$ and $\readCond{n,W,\var{base},\var{end}}$, then $\npair{(\perm, \var{base}, \var{end}, a)} \in \stder(W)$.
\end{lemma} %TODO: Change the condition here to be writeCondition (and readCondition)
\lau{With the current definition of the expression relation, an entry pointer will never be in the expression relation, but I guess we do not want that restriction. Say the adv capability in the ticket dispenser lemma, we want that to be an entry capability.}
\lau{In the ticket dispenser lemma, we need to make some assumptions about the adversary to be able to use the FTLR. What should these assumptions be? So far I have the assumption that it is all code (no capabilities), but I do not see how I conclude the write condition from this.}
\dominique{Something I find interesting here, is that the FTLR can be interpreted to say that x permissions have no real security value: if you have an arbitrary r or rw capability that is valid, then the FTLR says that it remains valid if we change it to rx or rwx (respectively).}
